# Binomial

A Bernoulli random variable has only two possible values: 1 and 0, with probabilities p and 1-p, respectively.
$$
P(X=x)=p(x)=
\begin{cases}
p^x(1-p)^{1-x}, \text{if x=0 or x=1}\\
0, \text{otherwise}\\
\end{cases}\\
EX = p \\
VarX = p(1-p)
$$

## Binomial Distribution

Suppose there are n identical, independent Bernoulli random variables: X1...Xn. Each $X_i$ has probability p of being 1. Then $Y=\int_{i=1}^{n}X_i$ follows the binomial(n,p) distribution, where $EY=np$ and $VarY = np(1-p)$
$$
P(Y=t) = \binom{N}{k}p^y(1-p)^{n-y}, y=0,...,n
$$

```{r}
n=30; p=1/2;
dbinom(0,n,p)
dbinom(15,n,p)
dbinom(30,n,p)
```


```{r}
y0=15; sum(dbinom(0,y0,n,p));pbinom(y0,n,p)
1-pbinom(y0,n,p);pbinom(y0,n,p,lower.tail=F)
sum(dbinom((y0+1):n,n,p))
```

$$
X_i \text{iid from Bernoulli(p), i=1,...,n. }\\
H_0: p = p_0 vs. H_1: p>p_0\\
\text{Reject null if } Y \ge b_{alpha,p_0}\\
$$

$$
H_0: p = p_0 vs. H_1: p<p_0\\
\text{Reject null if Y} \le c_{alpha,p_0}\\
$$
$$
H_0: p = p_0 vs. H_1: p\ne p_0\\
\text{Reject null if Y} \ge b_{alpha,p_0}\\
\alpha_1 + \alpha_2 = \alpha
$$
```{r}
qbinom(0.95, size = 30, prob = 1/2);qbinom(0.05, 30,1/2, lower.tail =F)
pbinom(19,30,1/2);1-pbinom(19,30,1/2)
pbinom(19,30,1/2,lower.tail=F)
sum(dbinom(20:30,30,1/2))

```

$H_0: p=1/2$ vs.  $H_1: p\ne 1/2$

```{r}

n=30; y=23
2*pbinom(22,30,1/2, lower.tail=F)
binom.test(23,30,0.5)
```

## Normal approximation to the binomial distribution

* The approximation is best when the binomial distribution is symemtric- that is, when p =1/2
* A frequently used rule of thumb is that the approximation is reasonable when np > 5 and n(1-p) >5
* The approximation is especially useful for large values of n

## Central Limit Theorum

Let ${X_1,...,X_n}$ be a sequence of iid random variables with mean $\mu$ and variance $\sigma^2 < \infty$. Let $\bar{X} = \frac{1}{n} \int_{-\infty}^{x} e^-\frac{y^2}{2}\frac{1}{\sqrt{2\pi\sigma}}$. That is, $\frac{\sqrt(n)(\bar{X}_n - \mu)}{\sigma}$ has a standard normal distribution.


$$
\frac{\sqrt(n)(\bar{X} - \mu)}{\sigma} = \frac{n(\bar{X}_n - \mu)}{\sqrt{n\sigma^2}} \\
= \frac{\sum_{i=1}^{n}X_i-n\mu}{\sqrt{Var(\sum_{i=1}^2) X_i}}\\
N(0,1)
$$

Suppose that a coin is tossed 100 times and lands heads up 60 times. The number of heads Y~binomial(n=100,p=1/2). We calculate $P(Y\ge60)$.

$$
EY = 50, VarY = np(1-p) = 25\\
P(Y\ge60) = P(\frac{Y-50}{5} \ge \frac{60-50}{5})\\
\approx 1- \phi(2)
$$

```{r}
1-pbinom(50,100,1/2)
pbinom(59,100,1/2, lower.tail=F)
pnorm(2,0,1, lower.tail=F)
```


* $X_i$ \text{iid from Bernoulli(p)} \\
* $\hat{p} = \bar{X} = \frac{1}{2} \sum_{i=1}^{n} X_i, E\hat{p} = p, Var{\hat{p}} = p(1-p)/n$\\
* $H_0 : p = p_0 vs. H1: p \ne p_0$ \\
* Under H_0, $T = \frac{\sqrt{n}(\hat{p-p_0})}{\sqrt(\hat{p}(1-\hat{p}))} \rightarrow N(0,1)$ as n $\rightarrow \infty$
* Under H_0, $T = \frac{\sqrt{n}(\hat{p-p_0})}{\sqrt(p_0(1-p_0))} \rightarrow N(0,1)$ as n $\rightarrow \infty$
* Reject $H_0$ if $|T| > z_{1-\frac{\alpha}{2}}$


```{r}
n=30; p0=0.5; y=23; xbar=y/n;
tstat1=sqrt(n)*(xbar-p0)/sqrt(xbar*(1-xbar))
tstat2=sqrt(n)*(xbar-p0)/sqrt(p0*(1-p0))
round(c(xbar,tstat1,tstat2^2,2*pnorm(abs(tstat1),lower.tail=F),
2*pnorm(abs(tstat2),lower.tail=F),pchisq(tstat2^2,df=1,lower.tail=F)),6)

prop.test(y,n,p0,correct=F)
```

* Don't reject $H_0$ if $|T| \le z_{1-\alpha/2}$

$$
1-\alpha = P(\frac{\sqrt{n}|\hat{p}-p_0}{\sqrt{\hat{p}(1-\hat{p})}} \le z_{1-\frac{\alpha}{2}} \\
= P(-z_{1-\frac{\alpha}{2}}\sqrt{\hat{p}(1-\hat{p})}) \le \sqrt{n}(\hat{p} - p_0) \le z_{1-\frac{\alpha}{2}}\sqrt(\hat{p}(1-\hat{p}))\\
= P(\hat{p} - z_{1-\frac{\alpha}{2}}\sqrt{\hat{p}(1-\hat{p}/n)} \le p_0 \le \hat{p} + z_{1-\frac{\alpha}{2}}\sqrt(\hat{p}(1-\hat{p})/n)
$$


```{r}

alpha = 0.05; qn = qnorm(1-alpha/2)
sx = sqrt(xbar*(1-xbar)/n)
xbar+c(-1,1)*qn*sx
```

```{r}

library(binom)
binom.confint(x=23,n=30)
```

## Bernoulli Power Calculations





