# Sign test and Wilcoxon signed rank


$$\begin{equation}
\text{Paired data} (X_i, Y_i), i = 1...,n\\
(X_i, Y_i) \text{independent of} (X_j, Y_j), i \ne j \\
X_i \sim (\mu_x, \sigma_x^2), Y_i \sim (\mu_x, \sigma_y^2)\\
Cov(X_i, Y_i) = \sigma_{xy} = \rho\sigma_x\sigma_y \\
D_i = X_i -Y_i \sim N(\delta = \mu_x - \mu_y, \sigma_x^2 + \sigma_y^2 -2\rho\sigma_x\sigma_y) \\ 
D_i = \bar{X} - \bar{Y} \sim N(\delta, \frac{1}{n}(\sigma_x^2 + \sigma_y^2 - 2\rho\sigma_x\sigma_y)) \\ 
\end{equation}$$
```{r}

n=30; x=rnorm(n,mean=1,sd=1); y=rnorm(n,mean=1.1,sd=1);
t.test(x,y,paired=T)
t.test(x-y)
```



## Sign test and Wilcoxon signed rank test

Nonparametric version of the paired or one sample t. test. Primary interest is centered on the location (median) of the population. Two scenarios: pared data, such as pretreatment and posttreatment where we are looking at shift in location due to treatment, or one sample data, where obserbations from a single population about whose location we wish to make inferences.

Assumptions:

* Let $Z_i = Y_i - X_i, i =1,...,n$ The differences $Z_i,...Z_n$ are mutually independent. $X_i$ and $Y_i$ can be dependent
* Each $Z_i$ comes from a continuous population (not necessarily the same one) and has a common median $\theta$
* $P(Z_i \le \theta) = P(Z_i > \theta) = 1/2$
* $P(Z_i - \theta \le 0) = P(Z_i -0 > 0) = 1/2$
* $\theta$ is the treatment effect

Under $H_0$:
* Each of the distributions for the differences has median 0, corresponding to no shift in location due to the treatment. 
* Sign statistic is the number of positive $Z_i$'s...
* $T = \sum_{i=1}^{n}l_{{z_{i}>0}}$
* random variable $l_{{z_{i}>0}}$ follows a Bernoulli distribution with $p = \frac{1}{2}$
* T follows a binomial(n,1/2) distribution


## Example

```{r}

library(BSDA)
x=c(1.83,0.50,1.62,2.48,1.68,1.88,1.55,3.06,1.30);
y=c(0.878,0.647,0.598,2.05,1.06,1.29,1.06,3.14,1.29);
plot(x,y,las=1); abline(0,1);
z=x-y; z; sign(z) ; stat=sum(x>y);
stat; median(z);
2*pbinom(stat-1,length(z),1/2,lower.tail=F)
2*pbinom(stat-1,length(z),1/2,lower.tail=F)
SIGN.test(z);
```
* When $\theta = 0$, the distribution of the statistic T is symmetric about its mean $n/2$
* A natural estimator of $\theta$ is the amount that $\hat\theta$ should be subtracted from each $Z_i$ so that the value of Y, when applied to the shifted sample $Z_i - \hat\theta, ..., Z_n = \hat\theta$ is as close to n/2 as possible
* We estimate $\theta$ by the amount $\hat\theta$ that the Z sample should be shifted in order that $Z_i - \hat\theta, ..., Z_n = \hat\theta$ appears as a sample from a population with median 0

```{r}
z-median(z)
median(z-median(z))
```


## Large sample test

## Wilcoxon signed rank test

```{r}
n=10; rn=(n*(n+1)/2); tv=0:rn;
plot(tv,dsignrank(tv,n),type="h",ylab="",main=paste("n=",n),las=1);
x0=11;
psignrank(x0,n);
1-psignrank(rn-x0-1,n);
psignrank(rn-x0-1,n,lower.tail=F);
```

## Example

```{r}
rbind(z, rank(z)); n=length(z)
Tplusv=sign(z)*rank(abs(z));
tplus=sum(Tplusv[Tplusv>0]); tplus
psignrank(tplus-1,n,lower.tail=F)
wilcox.test(x, y, paired = T, alternative = "greater",conf.int=T);
zz=outer(z,z,"+"); lzz=zz[lower.tri(zz,diag=T)]; median(lzz)/2

median(z-0.46)
wilcox.test(z, alternative = "greater",conf.int=T)
```



## Large sample approximation

```{r}
wilcox.test(z, alternative = "greater", conf.int=T,
exact = FALSE, correct = F) ;
```


