# Density
### Parametric continuous distribution

* The sampled data are from a continuous distribution with density function $f_{\theta}$
* $f(x) \ge 0$ for all x
* $\int_{-\infty}^{\infty} f(x)dx = 1$
* $F(b) = P(X \le b) = \int_{-\infty}^{b}f(x)dx$
* $P(a < X \le b) = \int_{a}^{b} f(x)dx = F(b) - F(a), a \le b$

```{r}
n=50; xis=rnorm(n,0,1); xbar=mean(xis);sdx=sd(xis)
xs=seq(min(xis),max(xis),length=100); nden=dnorm(xs)
dene=dnorm(xs,xbar,sdx); ry=range(nden,dene)
plot(xs,nden,las=1,type="l",ylim=ry,xlab="x",ylab="Normal density")
lines(xs,dene,col=2); legend("topleft",col=c(1,2),lty=1,bty="n",
legend=c("Standard normal","Estimated"))
```

### Histogram Version

```{r}
hist(xis,las=1,xlab="observed x",main="")
```

### Nonparametric density estimation

Nonparametric estimation of f does not assume ahead of time that the density belongs to a certain class or family of parametric distribution functions
$P(x- \frac{h}{2} < X \le n + \frac{h}{2}) = \int_{x-\frac{h}{2}}^{x + \frac{h}{2}} f(u)du \\ = F(x+\frac{h}{2}) - F(x-\frac{h}{2})$

If f is smooth and h is small,
$\int_{x-\frac{h}{2}}^{x + \frac{h}{2}} f(u)du \approx h f(x)$

A reasonable estimate of f(x) is then
$\hat{f}_{h}x = \frac{F(x+ \frac{h}{2}) - F(x - \frac{h}{2})}{h}$

### The empirical distribution function

* Observation: $x_1, ..., X_n$
* $F_{n}(x) = \frac{1}{n}(\# x_{i} \le x)$
* Ordered observation: $x_{(1)} \le x_{(2)} \le ... \le x_{(n)}$
* $F_{n}(x) = 0 \text{, if } x \le x_{(1)}$; 
* $F_{n}(x) = \frac{1}{n}\text{, if } x_{1} \le x < x_{(2)}$; 
* $F_{n}(x) = \frac{k}{n}\text{, if } x_{(k)} \le x < x_{(k+1)}$; 
* If there is a single observation with value $x_i$, $F_n$ has a jump of height $1/n$ at $x_i;
* If there are r observations with the same value $x_i$, $F_{n}$ has a jump of height $r/n$ at $x_i$
* $F_{n}(x)$ gives the proportion of the numbers less than or equal to x

### Example

```{r}
fhat=function(obx,x,h){
n=length(obx)
sdif=abs(obx-x)/h
fhx=sum(sdif<=1/2)/(n*h)
return(fhx)
}
xobs=c(1,1.3,1.6,1.8,2.4); x=1.7
fhat(xobs,x,0.5)
## [1] 0.8
fhat(xobs,x,1)

h=1; do=density(xobs,bw=h,kernel="r",n=5,cut=0);
do$x

do$y

do=density(xobs,bw=h/(2*sqrt(3)),kernel="r",n=5,cut=0);

h1=0.5; d1=density(xis,bw=h1,kernel="r"); h2=1.5; d2=density(xis,bw=h2,
kernel="r"); rx=range(xs,d1$x,d2$x); ry=range(nden,d1$y,d2$y)
plot(d1$x,d1$y,las=1,type="l",xlim=rx,ylim=ry,xlab="x",ylab="f(x)")
lines(xs,nden,col=2); lines(d2$x,d2$y,col=4); legend("topleft",
legend=c(paste("h=",h1),"normal",paste("h=",h2)),col=c(1,2,4),
lty=1,bty="n")

```

### Kernal function

A kernal function K is a function such that

* $K(u) \ge 0, -\infty < u <\infty$
* $K(-u) = K(u)$
* $\int_{-\infty}^{\infty} K(u)du = 1$
Uniform, rectangular, or box, kernal:
$$
K(u) = \begin{cases}
    1 , \frac{-1}{2} \le u < \frac{1}{2} \\
    0 , otherwise
\end{cases}
$$

* It takes on the value 1 whenever an observation $x_i$ is near x, $u = (x_{i} - x)\h$
* The notion of nearness is determined by the bandwidth h
* Taking the sum, the estimate at x is the number of data points $x_i$ near x multiplied by the factor $1\nh$

Given the data, the kernal density estimate is 

$\hat{f}x = \frac{1}{nh}\sum_{i=1}^{n} K(\frac{x-x_{i}}{h})$

* Triangle kernal:
* Epanechnikov Kernal:
* Gaussian Kernel

Kernel function provides a weighted value for the nubmer of $x_i$ near x. In particular, data values near x give more weight than those further away. A common choice of bandwidth is $h = 1.06n^{\frac{-1}{5}}min\{\sigma^2, IQR/1.34\} $

### Triangle Kernal

```{r}
u=seq(-1,1,length=100); plot(u,1-abs(u),las=1,ylab="",type="l")
lines(c(-1/2,1/2),c(1,1),col=2); abline(v=c(-1/2,1/2),col=2,lty=3)

fhat=function(obx,x,h){
n=length(obx)
sdif=abs(obx-x)/h
fhx=sum( (1-sdif) *( sdif<=1) )/(n*h)
return(fhx)
}
h=0.5; x=1.7
fhat(xobs,x,h)

dg=density(xis,kernel="gaussian"); dr=density(xis,kernel="r");
de=density(xis,kernel="epanechnikov"); rx=range(xs,dg$x,dr$x,de$x)
ry=range(nden,dg$y,dr$y,de$y); plot(dg$x,dg$y,las=1,type="l",xlim=rx,
ylim=ry,xlab="x",ylab="f(x)"); lines(xs,nden,ylim=ry,col=2);lines(dr$x,
dr$y,ylim=ry,col=4); lines(de$x,de$y,ylim=ry,col="darkgreen");
legend("topright",legend=c("gaussian kernel","rectangle kernel",
"epanechnikov kernel","normal density"), col=c(1,4,"darkgreen",2),
lty=1,bty="n")
```












