--- 
title: "NonParametric Stats"
author: "Katherine Wilson"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---

# Hypothesis Testing

$$\begin{equation}
X \sim N(\mu,\sigma^2), f(x) = \frac{1}{\sqrt{2\pi\sigma}}e-\frac{(x-\mu)^2}{2\sigma^2}, F(x) = \int_{-\infty}^{x} \frac{1}{\sqrt{2\pi\sigma}}e\frac{-(t-\mu)^2}{2\sigma^2}dt
\end{equation}$$

**Normal Distributions**

```{r}
a =0
b=1
lx=100
x=seq(-4,4,length=lx)
fx = dnorm(x,a,b)

bx = 1
dbx=dnorm(bx,a,b)
dbx

plot(x,fx, type = "l", xlab = "x", ylab = "f(x)", main = "Normal density functions",
     las= 1)
dist_2 <- dnorm(x, mean=0, sd=4)
dist_3 <- dnorm(x, mean=1, sd=1)
dist_4 <- dnorm(x, mean=1, sd =4)
lines(x, dist_2, type = "l", col = "red")
lines(x, dist_3, type = "l", col = "green")
lines(x, dist_4, type = "l", col = "blue")
```




$$\begin{equation}
F(x) = P(X\le x) = 1 - P(X>x)
\end{equation}$$

when lower.tail == TRUE, probabiltiies are p[X<x]

```{undefined eval=FALSE, include=TRUE}
## density under the curve
dnorm(x=0, mean=0, sd=1)
dnorm(0,0,2)
dnorm(0,1,1)
dnorm(0,1,2)
## probability
pnorm(0,0,1)
pnorm(0,0,2)
pnorm(0,1,1)
pnorm(0,1,2)

```



```{r}
a=0
b=1
lx=100
x=seq(-4,4,length=lx)
fx=dnorm(x,a,b)

bx=1
dbx=dnorm(bx,a,b)

plot(x,fx,type="l",lwd=2,xlab="x",ylab="Density",main="N(0,1)",las=1,xaxt="n")
segments(bx,0,bx,dbx,lty=3)
sx=seq(-4,bx,length=lx)
polygon(c(sx,rev(sx)),c(dnorm(sx,a,b),rep(0,lx)),density=2,border=NA,lty=3)

text(-0.5,0.1,expression(P(X<=z[1-alpha])==1-alpha))
abline(h=0,lty=3)
axis(1,at=bx,label=expression(z[1-alpha]))

sx=seq(bx,4,length=lx)
polygon(c(sx,rev(sx)),c(dnorm(sx,a,b),rep(0,lx)),density=3,border=NA,
angle=-45,lty=3)
text(2.5,0.1,expression(P(X>z[1-alpha])==alpha))

```


**Power**

Probability of rejecting $H_0$ when it is true is type I error (falsely reject), aka significance of the test aka $\alpha$. Probability of not reject H_0 when it is false is a type II error, $\beta$. Power of test is $1-\beta$.

**Confidence Intervals**

Confidence set with a confidence coefficient of $1-\alpha$ is the set.


**Parametric Methods**

Parametric distributions qith populations

$$\begin{equation}
N(\mu, \sigma^2),binomial(n,p)
\end{equation}$$

```{r}
n=30
x=rnorm(n,mean=0,sd=1)
sort(x)
xbar=mean(x)
sdx=sd(x)
```




**One sided H_0**

$$\begin{equation}
H_0 : \mu = \mu_0 vs. H_1: \mu > Mu_0\\
T = \frac{\sqrt(x-\mu_0)}{\sigma} \sim \text{t distribution with n-1 degrees of freedom} \\
\text{Reject null if t} > t_{n-1,1-\alpha} \\
\text{Type I error}: \alpha = P_{\mu_o}(T>t_{n-1,1-\alpha}) \\
\text{Acceptance region} : T \le t_{n-1,1-\alpha}\\
1- \alpha = P_{\mu_0}(\frac{\sqrt{n}(\bar{X}-\mu_0)}{\hat\sigma}) \le t_{n-1,1-\alpha}\\
= P_{\mu_0}(\bar{X} - t_{n-1,1-\alpha}\frac{\hat\sigma}{\sqrt{n}} \le \mu_0)
\end{equation}$$

Below, the critical values for alpha of 0.05 is 1.699. Given a sample mean of 1, and our critical test statistic of -6.07, probability in the lower tail is 0.99, so fail to accept the alternative that the true mean is greater than 1.
```{r}
alpha = 0.05
tq = qt(1-alpha, n-1)
c(qt(1-alpha, n-1), qt(alpha,n-1, lower.tail = F))
mu0=1
tstat=sqrt(n)*(xbar-mu0)/sdx;c(tstat,tq,pt(tstat,n-1,lower.tail=F))
xbar-tq*sdx/sqrt(n)
t.test(x,alternative="greater",mu=mu0)

```
Testing against sample mean of -1, then the tstat is 5.45, and probability in lower tail is <.05, so accept the alternative that true mean is greater than -1.
```{r}
mu0=-1; tstat=sqrt(n)*(xbar-mu0)/sdx
c(tstat,tq,pt(tstat,n-1, lower.tail=F))
xbar-tq*sdx/sqrt(n)
t.test(x,alternative="greater",mu=mu0)

```





**Power**

$$\begin{equation}
\text{Reject null if }\bar{X} > \mu_0 + t_{n-1,1-\alpha}\frac{\hat\sigma}{\sqrt{n}}\\
\text{Type II error: do not reject null when it is false} \\
\beta = P_{H_1}\text{do not reject null}\\
\text{Power: the probability of rejecting null when alt is true}\\
1-\beta = P_{\mu}(\bar{X} > \mu_0 + t_{n-1,1-\alpha}\frac{\hat\sigma}{\sqrt{n}}) \\
= P_{\mu}(\frac{\sqrt{n}(X-\mu)}{\hat\sigma} > \frac{\sqrt{n}(\mu_0 -\mu)}{\hat\sigma} + t_{n-1,1-\alpha}
\end{equation}$$


```{r}
alpha = 0.05
tq = qt(1-alpha, n-1)
mu0=1
mu=1.5
tv = sqrt(n)*(mu0-mu)/sdx+tq
pt(tv,n-1,lower.tail=F)
power.t.test(n,delta = abs(mu0-mu), sd=sdx,
             type = c("one.sample"), alternative = "one.sided")

```

$$\begin{equation}
t_{n-1,\beta} = \frac{\sqrt{n}(\mu_0 - \mu)}{\hat\sigma} + t_{n-1,1-\alpha}\\
n = [\frac{\hat\sigma(t_{n-1,\beta}-t_{n-1,1-\alpha})}{\mu_0-\mu}]^2
\end{equation}$$


```{r}
beta =0.1
tb = qt(beta, n-1)
(sdx*(tb-tq)/(mu0-mu))^2
power.t.test(power= 0.9, delta = abs(mu0-mu), sd = sdx,
             type = c("one.sample"), alternative = "one.sided")
```

**Two sided H1**



$$\begin{equation}
H_0 : \mu = \mu_0 vs. H_1: \mu > Mu_0\\
\text{Reject null if |T|} > t_{n-1,1-\frac{\alpha}{2}} \\
\text{Acceptance region} : \frac{\sqrt{n}\mid{\bar{X} - \mu_0}\mid}{\hat\sigma} \le t_{n-1,1-\frac{\alpha}{2}}\\
1- \alpha = P_{\mu_0}(\frac{\sqrt{n}(\bar{X}-\mu_0)}{\hat\sigma}) \le t_{n-1,1-\frac{\alpha}{2}}\\
= P_{\mu_0}(\bar{X} - t_{n-1,1-\frac{\alpha}{2}}\frac{\hat\sigma}{\sqrt{n}} \le \mu_0 \le \bar{X} + t_{n-1,1-\frac{\alpha}{2}}\frac{\hat\sigma}{\sqrt{n}}) \\
\text{Length of CI} : 2t_{n-1,1-\frac{\alpha}{2}}\frac{\hat\sigma}{\sqrt{n}}
\end{equation}$$


```{r}
alpha = 0.05
tq = qt(1-alpha/2, n-1)
mu0=0
tstat = sqrt(n)*(xbar-mu0)/sdx
c(tstat, tq,2*pt(abs(tstat),n-1, lower.tail =F))
xbar+c(-1,1)*tq*sdx/sqrt(n)
t.test(x,mu=mu0)

mu0=1
tstat=sqrt(n)*(xbar-mu0)/sdx
c(tstat,tq,2*pt(abs(tstat),n-1,lower.tail = F))
xbar + c(-1,1)*tq*sdx/sqrt(n)
t.test(x,mu=mu0)

```
**Power**

```{r}
alpha = 0.05; tq = qt(1-alpha/2,n-1);mu0=1;
mu=1.5;
tv0=sqrt(n)*(mu0-mu)/sdx;
pt(tv0+tq,n-1,lower.tail=F) + pt(tv0-tq,n-1);

power.t.test(n,delta=abs(mu0-mu), sd=sdx,
type = c("one.sample"),alternative="two.sided")

power.t.test(n,delta=abs(mu0-mu),sd=sdx,
             type =c("one.sample"),alternative="two.sided",strict=T)
```


