--- 
title: "NonParametric Stats"
author: "Katherine Wilson"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---

# Hypothesis Testing

$$\begin{equation}
X \sim N(\mu,\sigma^2), f(x) = \frac{1}{\sqrt{2\pi\sigma}}e-\frac{(x-\mu)^2}{2\sigma^2}, F(x) = \int_{-\infty}^{x} \frac{1}{\sqrt{2\pi\sigma}}e\frac{-(t-\mu)^2}{2\sigma^2}dt
\end{equation}$$

**Normal Distributions**

```{r}
a =0
b=1
lx=100
x=seq(-4,4,length=lx)
fx = dnorm(x,a,b)
fx

bx = 1
dbx=dnorm(bx,a,b)
dbx

plot(x,fx, type = "l", xlab = "x", ylab = "f(x)", main = "Normal density functions",
     las= 1)
dist_2 <- dnorm(x, mean=0, sd=4)
dist_3 <- dnorm(x, mean=1, sd=1)
dist_4 <- dnorm(x, mean=1, sd =4)
lines(x, dist_2, type = "l", col = "red")
lines(x, dist_3, type = "l", col = "green")
lines(x, dist_4, type = "l", col = "blue")
```




$$\begin{equation}
F(x) = P(X\le x) = 1 - P(X>x)
\end{equation}$$

when lower.tail == TRUE, probabiltiies are p[X<x]

```{undefined eval=FALSE, include=TRUE}
## density under the curve
dnorm(x=0, mean=0, sd=1)
dnorm(0,0,2)
dnorm(0,1,1)
dnorm(0,1,2)
## probability
pnorm(0,0,1)
pnorm(0,0,2)
pnorm(0,1,1)
pnorm(0,1,2)

```



```{r}
a=0
b=1
lx=100
x=seq(-4,4,length=lx)
fx=dnorm(x,a,b)

bx=1
dbx=dnorm(bx,a,b)

plot(x,fx,type="l",lwd=2,xlab="x",ylab="Density",main="N(0,1)",las=1,xaxt="n")
segments(bx,0,bx,dbx,lty=3)
sx=seq(-4,bx,length=lx)
polygon(c(sx,rev(sx)),c(dnorm(sx,a,b),rep(0,lx)),density=2,border=NA,lty=3)

text(-0.5,0.1,expression(P(X<=z[1-alpha])==1-alpha))
abline(h=0,lty=3)
axis(1,at=bx,label=expression(z[1-alpha]))

sx=seq(bx,4,length=lx)
polygon(c(sx,rev(sx)),c(dnorm(sx,a,b),rep(0,lx)),density=3,border=NA,
angle=-45,lty=3)
text(2.5,0.1,expression(P(X>z[1-alpha])==alpha))

```


**Power**

Probability of rejecting $H_0$ when it is true is type I error (falsely reject), aka significance of the test aka $\alpha$. Probability of not reject H_0 when it is false is a type II error, $\beta$. Power of test is $1-\beta$.

**Confidence Intervals**

Confidence set with a confidence coefficient of $1-\alpha$ is the set.


**Parametric Methods**

Parametric distributions qith populations

$$\begin{equation}
N(\mu, \sigma^2),binomial(n,p)
\end{equation}$$

```{r}
n=30
x=rnorm(n,mean=0,sd=1)
sort(x)
xbar=mean(x)
sdx=sd(x)
```

**One sided H_0**

$$\begin{equation}
H_0 : \mu = \mu_0 vs. H_1: \mu > Mu_0\\
T = \frac{\sqrt(x-\mu_0)}{\sigma} \sim t distribution with n-1 degrees of freedom \\
Reject H_0 if t > t_{n-1,1-\alpha} \\
Type I error: \alpha = P_{\mu_o}(T>t_{n-1,1-\alpha}) \\
Acceptance region: T \le t_{n-1,1-\alpha}\\
1- \alpha = P_{\mu_0}(\frac{\sqrt{n}(\bar{X}-\mu_0)}{\hat\sigma}) \le t_{n-1,1-\alpha}\\
= P_{\mu_0}(\bar{X} - t_{n-1,1-\alpha}\frac{\hat\sigma}{\sqrt{n}} \le \mu_0)
\end{equation}$$

